\subsection{Creational Patterns}

\frame{\tableofcontents[currentsection, currentsubsection]}

\begin{frame}[t]{Factory method}
	\begin{quote}
		If there are multiple ways to instantiate objects of your class, use \hhl{factory methods} rather than adding too much logic to the default constructor.
	\end{quote}

	\medskip
	\only<2| handout: 1>{
		\stress{\badc{Bad:}}
		\inputminted{python}{code/patterns/creational/uncertainties_bad.py}
	}
	\only<3-4| handout: 2>{
		\stress{\badc{Good:}}
		\inputminted[]{python}{code/patterns/creational/uncertainties_good.py}
	}
	\only<4| handout: 2>{
	Alternatively, you can also have subclasses that provide (implementations to) factory methods.
	}
\end{frame}

\begin{frame}[t]{Builder Pattern}
	\begin{quote}
		If you build a very complex class, try to instantiate (build) it in several steps.
	\end{quote}

	\stress{\badc{Bad:}}
	\inputminted{python}{code/patterns/creational/builder_bad.py}
	\only<2>{You will probably consider different fits and plots; \hhl{violates Single Responsibility Principle} $\lra$ Rather have \texttt{Fit} and \texttt{Plot} classes}
\end{frame}

%\begin{frame}[t]{Builder Pattern}
%	\begin{quote}
%		If you build a very complex class, try to instantiate (build) it in several steps.
%	\end{quote}
%
%	\stress{\badc{Bad:}}
%	\inputminted{python}{code/patterns/creational/builder_bad_2.py}
%
%\end{frame}

\begin{frame}[t]{Builder Pattern}
	\begin{quote}
		If you build a very complex class, try to instantiate (build) it in several steps.
	\end{quote}

	\stress{\goodc{Better:}}
	\inputminted[]{python}{code/patterns/creational/builder_good.py}
	\only<2>{
		\begin{itemize}
			\item What if we have multiple ways to build of the object?
			\item Do I want to have the \texttt{add\_mc\_component} method after I start using the data?
		\end{itemize}
		$\lra$ Have a separate \texttt{Data} and \texttt{Builder} hierarchy.
	}
%	\only<6| handout: 3>{
%		\stress{\goodc{Best:}}
%		\inputminted{python}{code/patterns/creational/builder_best.py}
%	}
\end{frame}

\begin{frame}[t]{Builder Pattern}
	\begin{quote}
		If you build a very complex class, try to instantiate (build) it in several steps.
	\end{quote}

	\stress{\goodc{Best:}}
	\inputminted{python}{code/patterns/creational/builder_best.py}
	And of course I could now create \texttt{AbstractData} and \texttt{AbstractBuilder} etc.
\end{frame}
